#440 Reproducibility and Statistics Assignment Homework

<!-- This is an Rmd comment. -->
<!--
You can make comments several lines long by inserting
the text in between the the start and the end of the 
comment.

Comments are where I'll tell you what to do. You should
use the space directly below the comments to do what
the comment asks. You should also describe what you're
doing to make it obvious to the reader so they can
easily follow your logic in a fully reproducible way.
For any statistical tests you run, you should explain
the results in plain English and present the statistics
as you would in an APA style manuscript. For any symbols
you don't know how to use/can't use, feel free to use the name
instead (e.g. alpha)
--> 

<!-- Load the politics.csv data file. -->

```{r}
politics<-read.csv("politics.csv")
```
I loaded the politics.csv data file.

<!--
Check the structure.
subject, party, testtime, minwage, and sex need to be factors.
optimismscore and income should be integer and numeric vectors.
Fix any variables that aren't right.
-->

```{r}
str(politics)```

I looked at the data.
  
```{r}politics$subject<-factor(politics$subject)```

I changed subject into a factor variable.
  
```{r}politics$testtime<-factor(politics$testtime, levels=c("pre", "post"))```

I refactored the variables "pre" and "post" so that "pre" precedes "post".

<!-- Find the minimum, mean, and variance for posttest
optimism scores. -->

```{r}
summary(politics$income)
```
I found the minimum, mean, and median incomes, but I want the minimun, mean, and variance for posttest optimism scores.

```{r}summary(politics$optimismscore[politics$testtime=="post"])```

I found the minimum, mean, and median for posttest optimisim scores.

```{r}var(politics$optimismscore[politics$testtime=="post"])```

I found the variance for posttest optimism scores.

<!-- Create a histogram of posttest optimism scores.
Feel free to use the default ugly settings since
we didn't cover how to do this in ggplot2. If
you're a real go-getter, see if you can figure
out how to make a prettier histogram with better
labels, etc. using ggplot2. -->

```{r}
hist(politics$optimismscore[politics$testtime=="post"])
```
I created a histogram ofposttest optimism scores.

<!-- Run a test to determine whether political affiliation
is independent of one's gender.
Present the data supporting your analyses in graphical or
tabular form.
-->

```{r}
tab<-table(politics$party[politics$testtime=="post"],politics$sex[politics$testtime=="post"])
tab
```
I used a table to calulate the frequency that individuals appear in different political groups. I focused on the posttest data to avoid counting people twice. 

```{r}
chisq.test(politics$party[politics$testtime=="post"],politics$sex[politics$testtime=="post"])
```

I used a Chi-Square test of independence to test the hypothesis that affiliation and support are independent. I focused on the posttest data to avoid counting people twice.Political affiliation is independent of gender, Chi-Square [2]= 0.73, p-value = 0.70.  

<!-- Run a test to determine whether or not males and females
have different incomes.
Create a graph to display your data that includes the relevant
means and standard errors.-->

```{r} t.test(politics$income[politics$sex=="male" & politics$testtime=="post"],politics$income[politics$sex=="female" & politics$testtime=="post"])```

To determine whether or not males and females have different posttest incomes, I ran an independent t-test. Males and females have different incomes, t(61.6) = -1.5714, p-value =
0.12

```{r}
library("dplyr")
temp<-politics[politics$testtime=="post",]%>%group_by(sex)%>%summarize(means=mean(income),
              sems=sd(income)/sqrt(length(income)))
library("ggplot2")
f<-ggplot(temp,aes(x=factor(sex),y=means))+
    geom_bar(stat="identity", color="black",fill=c("deeppink","deepskyblue2"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f
```

I created a figure by using the dplyr library.

```{r}
f<-f+ggtitle("Income by Sex")+
    labs(x="Sex", y="Posttest Income\n(higher=more income")+
    scale_x_discrete(breaks=c("female","male"),labels=c("Female","Male"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())
f
```
I made the graph more aesthetically pleasing and informative by changing the axis labels and text. 



<!-- See if male and females with different political affiliations
have different posttest optimism scores. Create a graph to display
your data that includes the relevant means and standard errors.-->

```{r}
 summary(aov(optimismscore~party*sex,data=politics[politics$testtime=="post",]))
```

I used a 2-way between-subjects ANOVA to seef if party affliliation and sex predict posttest optimism scores independently or in an interaction. The p-values for sex           0.843 and party:sex 0.304 do not tell me much in particular.  Maybe there's nothing going on, or maybe I just can't see it. The super tiny P- value for Party 4.2e-09 is definitely something. The two party columns have meaningfully different averages.


```{r}
temp<-politics[politics$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Posttest Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Posttest Optimism Scores")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

I created a plot via gplots library.






<!-- Create a line graph for the same data. With
error bars of course!-->

```{r}
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Posttest Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Posttest Optimism Scores")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

I created a line graph for the same data with error bars.



<!-- Run a test to see if optimism scores pre- and post
watching videos depends on sex.-->

```{r}
summary(aov(optimismscore~testtime*sex+Error(subject/testtime),data=politics))
```

To see if optimism scores before and after watching videos vary depending on sex, I ran a
2-way mixed ANOVA. I used a mixed ANOVA because the same
people indicated their optimism before and after the videos
(i.e. a within-subjects factor) and because there are different genders affiliated with different optimism scores (i.e.a between-subjects factor). 
The testtime p-value is 1.87e-08. Since the testtime groups have different averages, it indicates that optimism scores varied. 

<!-- Run a Multiple Regression Analysis to see
whether you can predict someones posttest optimism
scores from their pretest optimism scores and party
affiliation. Create a figure that shows the distributions
of pre- and posttest optimism scores and sex and create
lines that best fit the data for both sexes regardless of
whether the lines are warranted based on the statistical
analyses-->

```{r}
summary(lm(optimismscore~politics$optimismscore[politics$testtime=="pre"]+party,data=politics[politics$testtime=="post",]))```

I used Multiple Regression to predict posttime optimism scores from pretest optimism scores and party affiliation.
The Multiple R-squared=0.90, F-statistic=180.3 on 3 and 62 DF, and the p-value= 2.2e-16 indicate a lack of significance. 


```{r}
summary(lm(optimismscore~OptScorePre+sex,data=politics[politics$testtime=="post",]))
```


```{r}
OptScorePre<-politics$optimismscore[politics$testtime=="pre"]
f<-ggplot(politics[politics$testtime=="post",],aes(x=OptScorePre,y=optimismscore,color=sex))+
    geom_point(size=2)+
    geom_abline(intercept=8.44+0.22/2, slope=0.93,color=col1)+
    geom_abline(intercept=8.44-0.22/2, slope=0.93,color=col2)+
    scale_color_manual(values=c(col1,col2),breaks=c("male","female"),labels=c("Male","Female"))+
    ggtitle("Posttest Optimism by Pretest Optimism and Sex")+
    labs(x="Pretest Optimism Score\n(Higher=More)",y="Posttest Optimism Score\n(Higher=More)")+
    theme(plot.title=element_text(size=15,face="bold", vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold", vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold", vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.minor.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.2,.86))+
    theme(legend.background=element_blank())+
    theme(legend.title=element_blank())+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

I ran a regression using sex instead of party, so that I can find intercepts and create the plot as instructed. I created a scatter plot. The Multiple R-squared=0.90, F-statistic=269.6 on 2 and 63 DF, and the p-value= 2.2e-16 indicate a lack of significance. 

***Fin***
